---
title: "Writing images"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# use a temp directory so we can write logs as we wish
temp <- tempfile("logdir")
dir.create(temp)
knitr::opts_knit$set(root.dir = temp)
```

```{r setup}
library(tfevents)
```

tfevents has the ability to write log images so they can be
displayed in the TensorBoard dashboard. This guide describes
different ways you can log images with tfevents and how to
extend this functionality to other kinds of R objects that 
you might want to log as images in the dashboard.

## From arrays

The most common way to write images is when you have a batch of image data in 
an R array. For example, suppose we want to write the first 5 training samples
available in the Fashion MNIST dataset builtin into Keras.

```{r}
fashion_mnist <- keras::dataset_fashion_mnist()
train_imgs <- fashion_mnist$train$x[1:5,,]/255
str(train_imgs)
```

tfevents supports writing a batch of images from an array as long as the array has
4 dimensions: `batch_size`, `height`, `width` and `channels`. Thus for the images
in the Fashion MNIST dataset we need to add an extra dimension denoting that those
images have only one channel, ie. they are grayscale images.

```{r}
dim(train_imgs) <- c(dim(train_imgs), 1)
```

We can now log them to the default log directory (`logs`) with:

```{r}
log_event("Training samples" = summary_image(train_imgs))
```

If we open the TensorBoard UI in the logdir we will see something like the below:

```{r}
tensorflow::tensorboard(normalizePath("logs"), port = 6060)
```

Notice that images are also associated to a global step counter. If you save log
images at multiple steps you will be able to visualize hwo they change using
the slider in TensorBoard.

```{r batch, echo = FALSE, out.width="100%"}
webshot2::webshot("http://127.0.0.1:6060/?darkMode=false", vwidth = 1200)
```

## From ggplot objects

You can also write images from ggplot2 plots, for example:

```{r}
library(ggplot2)
p <- ggplot(mtcars, aes(x = hp, y = mpg)) + 
  geom_point()
```

You can log plots using `log_event` and `summary_image` and they will be displayed
in TensorBoard. The default size is 480x480 pixels but you can change the default by
passing `width` and `height` to `summary_image`:

```{r}
temp <- tempfile("logdir") # use another logdir 
set_default_logdir(temp)

log_event("Ggplot2 plot" = summary_image(p))
```

We now launch TensorBoard and see:

```{r}
tensorflow::tensorboard(temp, port = 6061)
```

```{r ggpl, echo = FALSE, out.width="100%"}
webshot2::webshot("http://127.0.0.1:6061/?darkMode=false", vwidth = 1200)
```

## Extending

Currently those are the only two builtin supported objects for writing images
with tfevents. However we provide lower level interfaces that allow you to extend
support for your own data type.

The `summary_image` function is an S3 generic, so you can implement it for your 
own type. For example, we could implement support for logging a raster image by
implementing a method for it. We also provide methods for lower level representation
of images, for example, `summary_image.raw` can write a `raw()` vector containing
a PNG encoded image.

Let's implement a method for writing raster images, created with `as.raster()`.
For example, we can use the first image in the Fashion MNIST dataset that we used
earlier.

```{r}
img <- as.raster(train_imgs[1,,,])
class(img)
```

Now let's define the `summary_image` method for the `raster` class. We can get a PNG encoded image by saving it to a file using `grDevices::png` and then reading it back 
to R.

Note that it's important to have `metadata` and `tag` arguments in the signature
so your users can customize hwo the image is displayed in TensorBoard.

```{r}
summary_image.raster <- function(img, ..., width = 480, height = 480, metadata = NULL, tag = NA) {
  temp <- tempfile(fileext = ".png")
  on.exit({unlink(temp)}, add = TRUE)

  grDevices::png(filename = temp, width = width, height = height, ...)
  plot(img)
  dev.off()

  sze <- fs::file_info(temp)$size
  raw <- readBin(temp, n = sze, what = "raw")

  summary_image(
    raw,
    width = width,
    height = height,
    colorspace = 4, # number of channels in the IMG. since PNG - 4
    metadata = metadata,
    tag = tag
  )
}
```

We can now save the raster image with:

```{r}
temp <- tempfile("logdir") # use another logdir 
set_default_logdir(temp)

log_event("Raster plot" = summary_image(img))
```

It will be displayed in the dashboard:

```{r raster, echo = FALSE, out.width="100%"}
tensorflow::tensorboard(temp, port = 6062)
webshot2::webshot("http://127.0.0.1:6062/?darkMode=false", vwidth = 1200)
```



